Dark matter halo shapes at $z=0$ in the Auriga simulations

### 2019-08-03 Document/paper.tex
up

The 30 zoom-in halos have a dark matter particle mass of $\sim 3\times
10^5$\Msun while the barynic mass resolution is $\sim 5\times 10^4$\Msun.
The softening lenght for gravitational force computation for stellar
particles and high-resolution dark matter particles 
is fixed to be 500 $h^{-1}$ pc in comoving coordinates up to $z=1$,
and 396 pc in physical coordinates afterwards.
The gravitational softening length for gas cells changes with the mean
cell radius but is limited to be larger than the stellar softening
lenght and 500 $h^{-1}$ pc comoving. 
From now on we refer to the haloes simulated with baryonic physics as the
MHD sample and to the haloes simulated with dark matter only as the
DMO sample.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{delta_triaxiality_distro.pdf}
\caption{
  Cumulative distribution for the changes in triaxility in simulations
  with different physics, $\Delta T=T_{\rm MHD}-T_{\rm DMO}$.
  At the virial radius all halos become less triaxial in the MHD
  simulations, the change is stronger towards the halo center.
  Only a small fraction of the halos ($\sim 1/15$) presents the
  opposite trend of becoming more triaxial towards the center.}
\label{fig:delta_triaxial_cumulative}
\end{figure}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{correlations_twisting_triaxiality_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alignment (measured as the standard deviation of the alignment angles)
  as a function of the halo triaxialiy at the radius indicated
  in each panel's title.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient while the line shows the result of the
  minimum squares fit to the data.
  Twisting does not correlate with triaxility in the same way at every
  radii. The strongest correlation appears around $0.12R_{200}$ ($28\pm2$ kpc).} 
\label{fig:alignment_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disc.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disc properties. 
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best minimum squares fit to a line.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disc to total mass ratio and the last
  column includes the disc optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^{-2}$ \citep{auriga}.
  The largest correlations are found for the two smaller radii
  ($0.06R_{200}$ and $0.12R_{200}$).
  Extended and massive stellar discs with a low gas content at its
  core are correlated with low dark matter triaxilities.
  The correlation decreases as one approaches larger radii.}
\label{fig:disc_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the same
  baryonic disc properties in Figure \ref{fig:disc_correlations}.
  The label with the $\rho$ value corresponds to the Spearman's rank
  correlation coefficient.
  The line is the best minimum squares fit to a line.
  In this case the disc to total mass ratio does not correlate with the alignment twisting
  The gas density at the center and the disc optical radius still show
  a strong correlation, albeit weaker than what they had with the triaxility.}
\label{fig:twisting_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_ages_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the
  mean stellar age in the bulge/disc and the lookback formation time of the
  dark matter halo (time at which it has half of its present mass).
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient.
  The line is the best minimum squares fit to a line.}
\label{fig:twisting_age}
\end{figure*}

@@

Figure \ref{fig:resolution} shows the results for the six halos
available at the higher Level3 resolution and the corresponding
halos at our baseline Level4 resolution. 
The dotted lines compare the DMO simulations showing that down to
scales of $1$ kpc the two simulations give comparable results.
In MHD simulations the situation is different, as expected. 
With the exception of Au-27 the shape ratios $b/a$ and $c/a$ start
to distantiate from each other for radii smaller than $15$ kpc.
Above these radii, the differences between the triaxility values at
Level4 and Level3 in MHD simulations are small and never exceed $0.1$.}

For this reason, in order to make the DM halo shape results comparable
against stellar disc shape measurements, we report all our shape
results above $10$ kpc.

>>>>

The 30 zoom-in halos have a dark matter particle mass of $\sim 3\times
10^5$\Msun while the barynic mass resolution is $\sim 5\times 10^4$\Msun.
The softening lenght for gravitational force computation for stellar
particles and high-resolution dark matter particles 
is fixed to be 500 $h^{-1}$ pc in comoving coordinates up to $z=1$,
and 396 pc in physical coordinates afterwards.
The gravitational softening length for gas cells changes with the mean
cell radius but is limited to be larger than the stellar softening
lenght and 500 $h^{-1}$ pc comoving. 
{\bf We refer to this setup as the Level4 resolution. 
Six of these haloes were also simulated at a higher Level3 resolution
that represent an increase by a factor of 8 in mass and 2 in spatial
resolution.  
We use the higher resolution haloes to define a minimal radius above
which our results can be considered as converged.} 
From now on we refer to the haloes simulated with baryonic physics as the
MHD sample and to the haloes simulated with dark matter only as the
DMO sample.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{delta_triaxiality_distro.pdf}
\caption{
  Cumulative distribution for the changes in triaxility in simulations
  with different physics, $\Delta T=T_{\rm MHD}-T_{\rm DMO}$.
  At the virial radius all halos become less triaxial in the MHD
  simulations, the change is stronger towards the halo center.
  Only two haloes in the sample show the opposite trend of becoming
  more triaxial (but only at the smaller radii).} 
\label{fig:delta_triaxial_cumulative}
\end{figure}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{correlations_twisting_triaxiality_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alignment (measured as the standard deviation of the alignment angles)
  as a function of the halo triaxialiy at the radius indicated
  in each panel's title.
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and error bars estimated via
  jackknife resampling) while the line shows the result of the
  minimum squares fit to the data.
  Twisting does not correlate with triaxility in the same way at every
  radii. The strongest correlation appears around $0.12R_{200}$ ($28\pm2$ kpc).} 
\label{fig:alignment_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{correlation_T_MHD_disc.pdf}
\end{center}
\caption{Correlations between the halo triaxility at different radii
  and baryonic disc properties. 
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and error bars estimated via
  jackknife resampling).
  The line is the best minimum squares fit to a line.
  The x-axis in the first column is the gas density at the center of
  the galaxy with in a sphere of radius  $1$ kpc \citep{Pakmor17};
  the second column shows the disc to total mass ratio and the last
  column includes the disc optical radius defined to be the radius at which the
  $B$-band surface brightness drops below 25 mag arcsec$^{-2}$ \citep{auriga}.
  The largest correlations are found for the two smaller radii
  ($0.06R_{200}$ and $0.12R_{200}$).
  Extended and massive stellar discs with a low gas content at its
  core are correlated with low dark matter triaxilities.
  The correlation decreases as one approaches larger radii.}
\label{fig:disc_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the same
  baryonic disc properties in Figure \ref{fig:disc_correlations}.
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and error bars estimated via
  jackknife resampling).
  The line is the best minimum squares fit to a line.
  In this case the disc to total mass ratio does not correlate with the alignment twisting
  The gas density at the center and the disc optical radius still show
  a strong correlation, albeit weaker than what they had with the triaxility.}
\label{fig:twisting_correlations}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{correlations_ages_angles_alignment_MHD.pdf}
\end{center}
\caption{Twisting in the halo-disc alingment as a function of the
  mean stellar age in the bulge/disc and the lookback formation time of the
  dark matter halo (time at which it has half of its present mass).
  The label with the $\rho$ value corresponds to the Spearmans rank
  correlation coefficient (mean value and error bars estimated via
  jackknife resampling).
  The line is the best minimum squares fit to a line.}
\label{fig:twisting_age}
\end{figure*}

@@

Figure \ref{fig:resolution} shows the results for the six halos
available both at Level3 and Level4 resolutions.
The dotted lines compare the DMO simulations showing that down to
scales of $1$ kpc the two resolutions yield comparable results
\citep{VeraCiro11}. 
In MHD simulations the situation is different, as expected. 
With the exception of Au-27 the shape ratios $b/a$ and $c/a$ start
to differentiate from each other at radii smaller than $15$ kpc.
Above these radii, the differences between the triaxility values at
Level4 and Level3 in MHD simulations are small and never exceed $0.1$.
For this reason, in order to make the DM halo shape results comparable
in DMO and MHD simulations we only report measurements above $15$
kpc.}
